On the surprising effectiveness of a simple matrix exponential derivative approximation, with application to global SARS-CoV-2

### 2023-10-30 manuscript/revision_letter.tex
ready to ship
really ready to ship

\begin{reply}
This is a good question.  Figure \ref{fig:randEffs} show posterior means of the exponentiated random effects, which amount to multiplicative deviations from the fixed-effect model.  As one might expect from the Bayesian bridge prior with $\alpha=1/4$ we use (Figure \ref{fig:truthMean}), only a small number of random effects deviate from 0, but those that do deviate are highly suggestive.  We stop short of interpreting exactly why the transition from the US to Hubei, CN, exhibits largest deviation (a 1.81-fold increase).  We have added this figure and the following short discussion to the appendix:
	\begin{quote}
		Figure \ref{fig:randEffs} displays posterior means of the exponentiated random effects, which one may interpret as multiplicative deviations from the fixed effects' contributions to the generator matrix.  Whereas the vast majority of exponentiated random effects have posterior means close to 1, indicating no deviation from the fixed-effect model, a few exhibit posterior means that are significantly greater than 1.  In particular, the rate element corresponding to transfer from the US to Hubei, CN, exhibits a 1.81-fold random-effect derived increase to the fixed-effect component.  This large multiplicative increase agrees with, but goes beyond, the influence that the Hubei asymmetry holds for the entire generator matrix model. 
	\end{quote}	
\end{reply}

@@

Despite these theoretical guarantees, the method of \citet{magee2023random} performed even better than expected in terms of relatively high acceptance rates and relatively low autocorrelation between samples.  From our other previous work with surrogate-trajectory HMC \citep{li2019neural}, we knew this must suggest that the first-order matrix exponential gradient approximation must be pretty good.   
As a result, our goal in this paper was to find out what makes the first-order approximation so good and provide a salient example that illustrates the efficacy of the approximation in higher dimensions.

In short, our original goal was not to prove validity of surrogate-trajectory HMC, which we saw as already established through simulation studies, real-world analyses and measure theoretic arguments.

>>>>

\begin{reply}
This is a good question.  Figure \ref{fig:randEffs} shows posterior means of the exponentiated random effects, which amount to multiplicative deviations from the fixed-effect model.  As one might expect from the Bayesian bridge prior with $\alpha=1/4$ we use (Figure \ref{fig:truthMean}), only a small number of random effects deviate from 0, but those that do deviate are highly suggestive.  We stop short of interpreting exactly why the transition from the US to Hubei, CN, exhibits largest deviation (a 1.81-fold increase).  We have added this figure and the following short discussion to the appendix:
	\begin{quote}
		Figure \ref{fig:randEffs} displays posterior means of the exponentiated random effects, which one may interpret as multiplicative deviations from the fixed effects' contributions to the generator matrix.  Whereas the vast majority of exponentiated random effects have posterior means close to 1, indicating no deviation from the fixed-effect model, a few exhibit posterior means that are significantly greater than 1.  In particular, the rate element corresponding to transfer from the US to Hubei, CN, exhibits a 1.81-fold random-effect derived increase to the fixed-effect component.  This large multiplicative increase agrees with, but goes beyond, the influence that the Hubei asymmetry holds for the entire generator matrix model. 
	\end{quote}	
\end{reply}

@@

Despite these theoretical guarantees, the method of \citet{magee2023random} performed even better than expected in terms of relatively high acceptance rates and relatively low autocorrelation between samples.  From our other previous work with surrogate-trajectory HMC \citep{li2019neural}, we knew this must suggest that the first-order matrix exponential gradient approximation must be pretty good.

As a result, our goal in this paper was to find out what makes the first-order approximation so good and provide a salient example that illustrates the efficacy of the approximation in higher dimensions.  
In short, our original goal was not to prove validity of surrogate-trajectory HMC, which we saw as already established through simulation studies, real-world analyses and measure theoretic arguments.
